---
title: "Quantifying the Impact of Retrofits and Nature on Equity and Sustainability"
author: "Devan Cantrell Addison-Turner, Yingjie Li, Anthony Dylan Kinslow II, Gretchen Cara Daily, and Rishee Kumar Jain"
date: "2024-06-27"
output: html_document
---


# Setup

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


## Packages --------------------------------------------------------------------
# install.packages("readr")
# install.packages("tidyverse")
# install.packages("dplyr")

library(readr)
library(dplyr)
library(tidyverse)


### --- to impute data ----
# install.packages("mice")
# install.packages("lattice")
# install.packages("VIM")
library(mice)
library(lattice)
library(VIM) ## aggr()


### ---- Matching ----
#install.packages("MatchIt")
##Note: MatchIt package may not install or load on MacOS. Instead, use Windows OS.

library(MatchIt)


### --- viz ----
# install.packages("ggplot2")
library(ggplot2)
```

# Data

## Input data

### Raw data 

```{r}

#California, USA K-12 Schools Raw Dataset to study 3 interventions effects on
#on Chronic Absenteeism during the COVID-19 Pandemic

CASchools_PM_Data <- read_csv("./data/ChronicAbsenteeism_AllSchools_PMData_2.csv")

#changing variable to numeric
CASchools_PM_Data2 <- CASchools_PM_Data %>%
  dplyr::mutate(
    Percent_Water = as.numeric(Percent_Water), 
    Percent_Developed = as.numeric(Percent_Developed),
    Percent_Forest = as.numeric(Percent_Forest),
    Percent_Pasture = as.numeric(Percent_Pasture),
    Staff = as.numeric(Staff), 
    NDVI_median = as.numeric(NDVI_median),
    Land_Surface_Temp = as.numeric(Land_Surface_Temp),
    Year_built = as.numeric(Year_built),
    Median_income_BG = as.numeric(Median_income_BG),
    Percent_FRPM = as.numeric(Percent_FRPM),
    Students_with_Disabilities = as.numeric(Students_with_Disabilities),
    Asthma_Rates = as.numeric(Asthma_Rates),
    COVID_infections = as.numeric(COVID_infections),
    Pollution_Burden = as.numeric(Pollution_Burden),
    Chronic_Absenteeism = as.numeric(Chronic_Absenteeism),
    retrofit = as.numeric(retrofit),
    ndvi_median_binary = as.numeric(ndvi_median_binary),
    retrofit_and_ndvi = as.numeric(retrofit_and_ndvi),
    )  %>%
    dplyr::select(
      #removed variables we do not need
                -Percent_LandCoverTotal,
                -Diversity_Index,
                -Average_NDVI_median,
                -Student_Staff_Ratio,
                -StudentEnrollment_mean,
              ) %>%
  as.data.frame() %>%
#We moved the school ID as column as row name to keep all the data in numeric
  column_to_rownames(var = 'School_ID')
```


### Impute data

  All multiple imputation techniques start with the MAR assumption (Missing at Random). 

  https://library.virginia.edu/data/articles/getting-started-with-multiple-imputation-in-r

```{r}

# Visualize Proportion of missing data
VIM::aggr(CASchools_PM_Data2, cex.axis=.425)

#Number of rows that would not be removed due to missing data
nrow(na.omit(CASchools_PM_Data2))

# Multivariate Imputation by Chained-Equations (MICE)
#fill missing values in dataset by prediction through multiple iterations using
#the MICE algorithm
CASchools_PM_Dataimp <- mice(CASchools_PM_Data2)

# Complete dataset after imputation and store as a new dataframe
CASchools_PM_Datacomplete <- complete(CASchools_PM_Dataimp)

#Visualize Proportion of missing data in Imputed Dataset
aggr(CASchools_PM_Datacomplete, cex.axis=.425)

#Number of rows that would not be removed due to missing data
nrow(na.omit(CASchools_PM_Datacomplete))
```

### Data Normalization
```{r}

#Create a function to normalize variables on a scale from 0 to 1 then apply to dataset

maxmin_scalar <- function(x){
  (x- min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm = T))
  }
scaling_maxmin <- as.data.frame(lapply(CASchools_PM_Data2, maxmin_scalar))

scaling_maxmin$School_ID <- rownames(CASchools_PM_Data2)

scaling_maxmin_dropna <- scaling_maxmin %>%
  tidyr::drop_na()

######### Imputed Dataset

maxmin_scalar <- function(x){
  (x- min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm = T))
  }
scaling_maxmin_complete <- as.data.frame(lapply(CASchools_PM_Datacomplete, maxmin_scalar))

scaling_maxmin_complete$School_ID <- rownames(CASchools_PM_Datacomplete)
```




# Pair-Matching Algorithm - 3 Scenarios

## Using non-imputed data

### Scenario 1: Energy Efficiency (EE) Retrofits in K-12 Schools

```{r}
#
#1:1 optimal PS matching with optimal matching on School covariates
m.out1 <- matchit(retrofit ~ 
                    NDVI_median +	
                    Percent_Water	+ Percent_Developed + Percent_Forest +	
                    Percent_Pasture + Land_Surface_Temp + Year_built +
                    Median_income_BG +	Percent_FRPM + Staff +
                    Students_with_Disabilities +	Asthma_Rates +	
                    COVID_infections +	Pollution_Burden,
                  data = scaling_maxmin_dropna,
                  method = "optimal") 
summary(m.out1)


#Scenario 1: Energy Efficiency Retrofits in K-12 Schools
m.out1
summary(m.out1, un = FALSE)

m_data <- match.data(m.out1)
mm <- na.omit(m.out1$match.matrix)
m_data <- m_data[rownames(m_data) %in% rownames(mm) |
                   rownames(m_data) %in% mm,]
#print(m_data)
```



### Scenario 2: Nature Exposure in K-12 Schools
```{r}
m.out2 <- matchit(ndvi_median_binary ~ 
                    retrofit +	
                    Percent_Water	+ Percent_Developed + Percent_Forest +	
                    Percent_Pasture + Land_Surface_Temp + Year_built +
                    Median_income_BG +	Percent_FRPM + Staff +	
                    Students_with_Disabilities +	Asthma_Rates +	
                    COVID_infections +	Pollution_Burden,
                   data = scaling_maxmin_dropna,
                   method = "optimal")
summary(m.out2)

### Scenario 2: Nature Exposure in K-12 Schools
m.out2
summary(m.out2, un = FALSE)
m_data2 <- match.data(m.out2)
mm2 <- na.omit(m.out2$match.matrix)
m_data2 <- m_data2[rownames(m_data2) %in% rownames(mm2) |
                   rownames(m_data2) %in% mm2,]
```



### Scenario 3: Energy Efficiency Retrofits and Nature Exposure in K-12 Schools 
```{r}
m.out3 <- matchit(retrofit_and_ndvi ~ 
                    Percent_Water	+ Percent_Developed + Percent_Forest +	
                    Percent_Pasture + Land_Surface_Temp	+ Year_built +
                    Median_income_BG +	Percent_FRPM + Staff +
                    Students_with_Disabilities +	Asthma_Rates +	
                    COVID_infections +	Pollution_Burden,
                   data = scaling_maxmin_dropna,
                   method = "optimal")
summary(m.out3)


m.out3
summary(m.out3, un = FALSE)
m_data3 <- match.data(m.out3)
mm3 <- na.omit(m.out3$match.matrix)
m_data3 <- m_data3[rownames(m_data3) %in% rownames(mm3) |
                   rownames(m_data3) %in% mm3,]
```


## Using Imputed Data

### Scenario 1: Energy Efficiency Retrofits in K-12 Schools with Imputed Dataset

```{r}
#1:1 optimal PS matching with optimal matching on School covariates
m.out1_complete <- matchit(retrofit ~ NDVI_median +	Percent_Water	+
                           Percent_Developed + Percent_Forest +	
                           Percent_Pasture + Land_Surface_Temp + Year_built +
                           Median_income_BG +	Percent_FRPM + Staff +
                           Students_with_Disabilities +	Asthma_Rates +	
                           COVID_infections +	Pollution_Burden,
                  data = scaling_maxmin_complete,
                  method = "optimal") 
summary(m.out1_complete)
 
m.out1_complete
summary(m.out1_complete, un = FALSE)
m_data_complete <- match.data(m.out1_complete)
```


### Scenario 2: Nature in K-12 Schools with Imputed Dataset

```{r}
m.out2_complete <- matchit(ndvi_median_binary ~ retrofit +	Percent_Water	+
                           Percent_Developed + Percent_Forest +	
                           Percent_Pasture + Land_Surface_Temp + Year_built +
                           Median_income_BG +	Percent_FRPM + Staff +	
                           Students_with_Disabilities +	Asthma_Rates +	
                           COVID_infections +	Pollution_Burden,
                   data = scaling_maxmin_complete,
                   method = "optimal")
summary(m.out2_complete)
m.out2_complete
summary(m.out2_complete, un = FALSE)
m_data2_complete <- match.data(m.out2_complete)
```


### Scenario 3: Energy Efficiency Retrofits and Nature in K-12 Schools with Imputed Dataset

```{r}
m.out3_complete <- matchit(retrofit_and_ndvi ~ Percent_Water	+
                           Percent_Developed + Percent_Forest +	
                           Percent_Pasture + Land_Surface_Temp + Year_built +
                           Median_income_BG +	Percent_FRPM + Staff +
                           Students_with_Disabilities +	Asthma_Rates +	
                           COVID_infections +	Pollution_Burden,
                  data = scaling_maxmin_complete,
                  method = "optimal")
summary(m.out3_complete)

m.out3_complete
summary(m.out3_complete, un = FALSE)
m_data3_complete <- match.data(m.out3_complete)
```


## Control and Treatment subsets

### Scenario 1
```{r}
### Scenario 1: Energy Efficiency Retrofits in K-12 Schools

#Control Group 
m_data_c <- m_data %>%
  dplyr::filter(retrofit == 0 )

#Treatment Group
m_data_t <- m_data %>%
  dplyr::filter(retrofit == 1)


## left join raw outcome data
raw <- CASchools_PM_Data2 %>%
  dplyr::select(Chronic_Absenteeism) %>%
  dplyr::mutate(School_ID = rownames(.)) %>%
  dplyr::rename('Chronic_Absenteeism_raw' = 'Chronic_Absenteeism')


m_data_c_withRaw <- m_data_c %>%
  left_join(., raw, by = 'School_ID')


m_data_t_withRaw <- m_data_t %>%
  left_join(., raw, by = 'School_ID')

m_data_wide <- merge(
  x = m_data_c, 
  y = m_data_t,
  by = 'subclass'
)

m_data_wide_2 <- merge(
  x = m_data_c_withRaw, 
  y = m_data_t_withRaw,
  by = 'subclass'
)
```


### Scenario 2

```{r}
## Scenario 2: Nature Exposure in K-12 Schools

#Control Group 
m_data2_c <- m_data2 %>%
  dplyr::filter(ndvi_median_binary == 0 )

#Treatment Group
m_data2_t <- m_data2 %>%
  dplyr::filter(ndvi_median_binary == 1)


## left join raw outcome data
raw_2 <- CASchools_PM_Data2 %>%
  dplyr::select(Chronic_Absenteeism) %>%
  dplyr::mutate(School_ID = rownames(.)) %>%
  dplyr::rename('Chronic_Absenteeism_raw' = 'Chronic_Absenteeism')


m_data2_c_withRaw <- m_data2_c %>%
  left_join(., raw_2, by = 'School_ID')


m_data2_t_withRaw <- m_data2_t %>%
  left_join(., raw_2, by = 'School_ID')

m_data_wide_3 <- merge(
  x = m_data2_c, 
  y = m_data2_t,
  by = 'subclass'
)

m_data_wide_4 <- merge(
  x = m_data2_c_withRaw, 
  y = m_data2_t_withRaw,
  by = 'subclass'
)
```


### Scenario 3
```{r}
## Scenario 3: Energy Efficiency and Nature Exposure in K-12 Schools

#Control Group 
m_data3_c <- m_data3 %>%
  dplyr::filter(retrofit_and_ndvi == 0 )

#Treatment Group
m_data3_t <- m_data3 %>%
  dplyr::filter(retrofit_and_ndvi == 1)


## left join raw outcome data
raw_3 <- CASchools_PM_Data2 %>%
  dplyr::select(Chronic_Absenteeism) %>%
  dplyr::mutate(School_ID = rownames(.)) %>%
  dplyr::rename('Chronic_Absenteeism_raw' = 'Chronic_Absenteeism')


m_data3_c_withRaw <- m_data3_c %>%
  left_join(., raw_3, by = 'School_ID')


m_data3_t_withRaw <- m_data3_t %>%
  left_join(., raw_3, by = 'School_ID')

m_data_wide_5 <- merge(
  x = m_data3_c, 
  y = m_data3_t,
  by = 'subclass'
)

m_data_wide_6 <- merge(
  x = m_data3_c_withRaw, 
  y = m_data3_t_withRaw,
  by = 'subclass'
)
```

####################### Imputed Data

```{r}
### Scenario 1: Energy Efficiency Retrofits in K-12 Schools with Imputed Data
#Control Group 
m_data_c_complete <- m_data_complete %>%
  dplyr::filter(retrofit == 0)

#Treatment Group
m_data_t_complete <- m_data_complete %>%
  dplyr::filter(retrofit == 1)

## left join raw outcome data
raw_complete <- CASchools_PM_Datacomplete %>%
  dplyr::select(Chronic_Absenteeism) %>%
  dplyr::mutate(School_ID = rownames(.)) %>%
  dplyr::rename('Chronic_Absenteeism_raw' = 'Chronic_Absenteeism')

m_data_c_withRaw_complete <- m_data_c_complete %>%
  left_join(., raw_complete, by = 'School_ID')

m_data_t_withRaw_complete <- m_data_t_complete %>%
  left_join(., raw_complete, by = 'School_ID')

m_data_wide_complete <- merge(
  x = m_data_c_complete, 
  y = m_data_t_complete,
  by = 'subclass'
)

m_data_wide_2_complete <- merge(
  x = m_data_c_withRaw_complete, 
  y = m_data_t_withRaw_complete,
  by = 'subclass'
)
```

```{r}
### Scenario 2: Nature Exposure in K-12 Schools with Imputed Data

#Control Group 
m_data2_c_complete <- m_data2_complete %>%
  dplyr::filter(ndvi_median_binary == 0 )

#Treatment Group
m_data2_t_complete <- m_data2_complete %>%
  dplyr::filter(ndvi_median_binary == 1)

## left join raw outcome data
raw_2_complete <- CASchools_PM_Datacomplete %>%
  dplyr::select(Chronic_Absenteeism) %>%
  dplyr::mutate(School_ID = rownames(.)) %>%
  dplyr::rename('Chronic_Absenteeism_raw' = 'Chronic_Absenteeism')


m_data2_c_withRaw_complete <- m_data2_c_complete %>%
  left_join(., raw_2_complete, by = 'School_ID')

m_data2_t_withRaw_complete <- m_data2_t_complete %>%
  left_join(., raw_2_complete, by = 'School_ID')

m_data_wide_3_complete <- merge(
  x = m_data2_c_complete, 
  y = m_data2_t_complete,
  by = 'subclass'
)

m_data_wide_4_complete <- merge(
  x = m_data2_c_withRaw_complete, 
  y = m_data2_t_withRaw_complete,
  by = 'subclass'
)
```

```{r}
### Scenario 3: Energy Efficiency and Nature Exposure in K-12 Schools with Imputed Data

#Control Group 
m_data3_c_complete <- m_data3_complete %>%
  dplyr::filter(retrofit_and_ndvi == 0)

#Treatment Group
m_data3_t_complete <- m_data3_complete %>%
  dplyr::filter(retrofit_and_ndvi == 1)


## left join raw outcome data
raw_3_complete <- CASchools_PM_Datacomplete %>%
  dplyr::select(Chronic_Absenteeism) %>%
  dplyr::mutate(School_ID = rownames(.)) %>%
  dplyr::rename('Chronic_Absenteeism_raw' = 'Chronic_Absenteeism')


m_data3_c_withRaw_complete <- m_data3_c_complete %>%
  left_join(., raw_3_complete, by = 'School_ID')


m_data3_t_withRaw_complete <- m_data3_t_complete %>%
  left_join(., raw_3_complete, by = 'School_ID')

m_data_wide_5_complete <- merge(
  x = m_data3_c_complete, 
  y = m_data3_t_complete,
  by = 'subclass'
)

m_data_wide_6_complete <- merge(
  x = m_data3_c_withRaw_complete, 
  y = m_data3_t_withRaw_complete,
  by = 'subclass'
)
```


# Statistical Analysis 

## Compute T-Test

### Using Raw data
```{r}
## Scenario 1: Energy Efficiency Retrofits in K-12 Schools 
TTestS1 <- t.test(m_data_wide_2$Chronic_Absenteeism_raw.x, m_data_wide_2$Chronic_Absenteeism_raw.y, paired = TRUE, conf.level = 0.95)
print(TTestS1)

## Scenario 2: Nature Exposure in Schools
TTestS2 <- t.test(m_data_wide_4$Chronic_Absenteeism_raw.x, m_data_wide_4$Chronic_Absenteeism_raw.y, paired = TRUE, conf.level = 0.95)
print(TTestS2)

## Scenario 3: Energy Efficiency Retrofits in K-12 Schools and Nature Exposure in K-12 Schools
TTestS3 <- t.test(m_data_wide_6$Chronic_Absenteeism_raw.x,m_data_wide_6$Chronic_Absenteeism_raw.y, paired = TRUE, conf.level = 0.95)
print(TTestS3)
```


### Using Imputed Data

```{r}

## Scenario 1: Energy Efficiency in K-12 Schools with Imputed Data
TTestS1_complete <- t.test(m_data_wide_2_complete$Chronic_Absenteeism_raw.x, m_data_wide_2_complete$Chronic_Absenteeism_raw.y, paired = TRUE, conf.level = 0.95)
print(TTestS1_complete)

## Scenario 2: Nature Exposure in K-12 Schools with Imputed Data
TTestS2_complete <- t.test(m_data_wide_4_complete$Chronic_Absenteeism_raw.x, m_data_wide_4_complete$Chronic_Absenteeism_raw.y, paired = TRUE, conf.level = 0.95)
print(TTestS2_complete)

## Scenario 3: Energy Efficiency and Nature Exposure in K-12 Schools with Imputed Data
TTestS3_complete <- t.test(m_data_wide_6_complete$Chronic_Absenteeism_raw.x, m_data_wide_6_complete$Chronic_Absenteeism_raw.y, paired = TRUE, conf.level = 0.95)
print(TTestS3_complete)
```


## Cohen's D

### Using Raw data
```{r}
#install.packages("devtools")
#install.packages("effsize")
#compute Effect Size using Cohen's D
library(effsize)

## Scenario 1: Energy Efficiency Retrofits in K-12 Schools 
print('Scenario 1')
cohen.d(m_data_wide_2$Chronic_Absenteeism_raw.x, m_data_wide_2$Chronic_Absenteeism_raw.y)


## Scenario 2: Nature Exposure in K-12 Schools 
print('Scenario 2')
cohen.d(m_data_wide_4$Chronic_Absenteeism_raw.x, m_data_wide_4$Chronic_Absenteeism_raw.y)


## Scenario 3: Energy Efficiency Retrofits and Nature Exposure in K-12 Schools 
print('Scenario 3')
d3 <- cohen.d(m_data_wide_6$Chronic_Absenteeism_raw.x, m_data_wide_6$Chronic_Absenteeism.y)
d3

es3 <- d3$estimate
es3


mean_c <- mean(m_data_wide_6$Chronic_Absenteeism.x, na.rm = T)
mean_c
```


### Using Imputed Dataset

```{r}

## Scenario 1: Energy Efficiency Retrofits in K-12 Schools with Imputed Dataset
cat('\n Scenario 1 ------ \n')
cohen.d(m_data_wide_2_complete$Chronic_Absenteeism_raw.x, m_data_wide_2_complete$Chronic_Absenteeism_raw.y)

## Scenario 2: Nature Exposure in K-12 Schools with Imputed Dataset
cat('\n Scenario 2 ------ \n')
cohen.d(m_data_wide_4_complete$Chronic_Absenteeism_raw.x, m_data_wide_4_complete$Chronic_Absenteeism_raw.y)

## Scenario 3: Energy Efficiency Retrofits and Nature Exposure in K-12 Schools  with Imputed Dataset
cat('\n Scenario 3 ------ \n')
cohen.d(m_data_wide_6_complete$Chronic_Absenteeism_raw.x, m_data_wide_6_complete$Chronic_Absenteeism_raw.y)
```


## Graph Effect Size
```{r}

library(ggplot2)

#' Standardized Mean Difference (Cohen's d)
#' Here we use the effect size in scenario 3, a combined effect of 
#'  Energy Efficiency Retrofits and Nature Exposure
ES <- es3 


#' get mean2 depending on value of ES from 
#'  d = (u1 - u2)/sd
#'  es = (mean1 - mean2)/sd
mean2 = mean_c 


sd <- 1
mean1 <- ES*sd + mean2

# create x sequence
x <- seq(mean2 - 3*sd, mean1 + 3*sd, .01)

# generate normal dist #1
y1 <- dnorm(x, mean2, sd)
# put in data frame
df1 <- data.frame("x" = x, "y" = y1)

# generate normal dist #2
y2 <- dnorm(x, mean1, 1)
# put in data frame
df2 <- data.frame("x" = x, "y" = y2)

# get y values under overlap
y.poly <- pmin(y1,y2)
# put in data frame
poly <- data.frame("x" = x, "y" = y.poly)

#plot with ggplot2
ggplot(df1, aes(x,y, color="control")) + 
  geom_line(linewidth=1) +
  geom_line(data=df2, aes(color="treatment"),linewidth=1) +
  geom_vline(xintercept = mean1, linetype="dotted") + 
  geom_polygon(aes(color=NULL), data=poly, fill="red", alpha=I(4/10),
               show_guide = F) +
  geom_vline(xintercept = 1, linetype="dotted") +
  geom_vline(xintercept = mean1, linetype="dotted") +  
  ggtitle(paste0(
    "Impact of Energy Efficiency Retrofits and Nature Exposure on Chronic Absenteeism",
    "\nCohen's d = ",  ES, ' (Large Effect)')) +
  # # change colors and legend annotation
  scale_color_manual("Group",
            values= c("treatment" = "red","control" = "black")) +
  # # remove axis labels
  ylab("Mean") + xlab("Standard Deviation (σ)") +
  theme_bw()+
  theme(legend.position = c(0.90, 0.85), 
        text = element_text(size = 11),
        plot.title = element_text(size=11))

ggsave(filename = './figures/effect_size.png', plot = last_plot(), width = 6.5, height = 4, units = "in", dpi = 300)
```
